---
title: "Global Disease Burden Analysis"
author: "Elena" # AUTHOR NAMES HERE!
output: github_document
---

```{r include = F}
knitr::opts_chunk$set(echo = FALSE)
```

# Introduction

This document compares the Disability-Adjusted Life Year (DALY) diseases burden for three countries: COUNTRY 1, COUNTRY 2, and COUNTRY 3. We compare the burden due to communicable, maternal, neonatal, and nutritional diseases (CMNN), non-communicable diseases (NCDs), and overall disease burden for these countries. 

The source data is from the Institute for Health Metrics and Evaluation (IHME) Global Burden of Disease (GBD) study. 

```{r child='cmnn_burden.Rmd'}

# Communicable, Maternal, Neonatal, and Nutritional Disease Burden (CMNN)


# Load necessary packages
pacman::p_load(tidyverse, knitr, here)

# Import the communicable diseases data
data_cmnn <- read_csv(here("data", "burden-of-disease-cmnn.csv"))
data_cmnn
```

## Table of Estimates for CMNN Burden Over Time

```{r}


# Use kable() from the knitr package to render the table.

pivoted_data <- data_cmnn %>%
  filter(Entity %in% c("Afghanistan", "Brazil", "Slovenia")) %>% 
  select(Entity, Year, `DALYs (Disability-Adjusted Life Years) - Communicable, maternal, neonatal, and nutritional diseases - Sex: Both - Age: Age-standardized (Rate)`) %>%
  pivot_wider(names_from = Entity, values_from = `DALYs (Disability-Adjusted Life Years) - Communicable, maternal, neonatal, and nutritional diseases - Sex: Both - Age: Age-standardized (Rate)`, names_prefix = "DALYs_")

# Render the table using kable()
kable(pivoted_data)

```

## Plot Showing Trends in CMNN Burden Over Time

```{r}
# Here, plot a line graph showing the trend of DALY burden for the chosen countries over time. You can use ggplot2 to create this plot. Each country's line should be a different color.
ggplot(pivoted_data, aes(x = Year)) +
  geom_line(aes(y = DALYs_Afghanistan, color = "Afghanistan")) +
  geom_line(aes(y = DALYs_Brazil, color = "Brazil")) +
  geom_line(aes(y = DALYs_Slovenia, color = "Slovenia")) +
  labs(x = "Year", y = "DALYs", color = "Country") +
  scale_color_manual(values = c("Afghanistan" = "blue", "Brazil" = "red", "Slovenia" = "green")) +
  theme_minimal()

```

## Summary of CMNN Burden Findings

Provide a brief analysis based on the data presented in the table and chart. Highlight any significant findings or patterns. About 3 sentences.

Slovenia has the lowest DALY of the three countries and has been decreasing in a steady line over the past 30 years. Afghanistan is still the country with the highest DALY in 2020, but has improved the most in the last 30 years.



```

```{r child='ncd_burden.Rmd'}
```

```{r child='overall_burden.Rmd'}
```
